---
title: "How to Design and Pre-Register Experimental Designs with Baseline Data"
author: "Experiments in Governance and Politics (EGAP)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Pre-Register a Simple Experimental Design using the registration package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates how to design and pre-register an experiment after baseline data has been collected.

```{r, echo = FALSE}
library(registration)
```

```{r, echo = FALSE}
## Note in this snippet we create the data that will be used as if it was "user data".

setwd(tempdir())

smp <- declare_sample_frame(
  individuals = list(
    income = declare_variable()),
  villages = list(
    development_level = declare_variable(multinomial_probabilities = 1:5/sum(1:5))
  ),
  lower_units_per_level = list(
    individuals = rep(1,1000), 
    villages = rep(5,200)
  ))

user_data          <- make_data(sample_frame = smp)

save(user_data, file = "baseline_data.RData")
```

First, we load the baseline data created by the user, and then define a set of covariates that will be simulated to conduct power analysis and for simulated analyses.

```{r covariates}
load("baseline_data.RData")

sample_frame <- declare_sample_frame(data = user_data, resample = TRUE, N = 250)

smp2 <- declare_sample_frame(
  individuals = list(
    income = declare_variable()),
  villages = list(
    development_level = declare_variable(multinomial_probabilities = 1:5/sum(1:5))
  ),
  N_per_level = c(50, 10),
  data = user_data, resample = TRUE)
```
  
Second, we define the potential outcomes, which will be simulated based on the baseline covariate data.

```{r}
potential_outcomes     <-  declare_potential_outcomes(
  condition_names = c("Z0","Z1"),
  outcome_formula = Y ~ .01 + 0*Z0 + .2*Z1 + .1*income,
  cluster_variable = "villages_id",
  ICC = .2
)
```

Third, we make a simulated data set and then declare the design.

```{r}
mock          <- make_data(sample_frame = sample_frame, potential_outcomes = potential_outcomes)
```



```{r}
design <- declare_design(potential_outcomes = potential_outcomes)
```



Fourth, we define one or more analyses we will run based on simulated data. This analysis will also be used for power analysis.


```{r}
analysis      <- declare_analysis(formula = Y ~ Z + income + development_level, treatment_variable = "Z",
                                  method = "lm")
```

Finally, we use the pre_register function to create the pre-registration document.

```{r}
pre_register(design = design, sample_frame = cov_object, 
             potential_outcomes = po_object, analysis = analysis, 
             registration_title = "Lady Tasting Tea", 
             registration_authors = c("Ronald A. Fisher"), 
             registration_abstract = "Description of the lady tasting tea experiment",
             temp_dir = TRUE)
```
