---
title: "How to Design and Pre-Register Experimental Designs with Baseline Data"
author: "Experiments in Governance and Politics (EGAP)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Pre-Register a Simple Experimental Design using the registration package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates how to design and pre-register an experiment after baseline data has been collected.

```{r, echo = FALSE}
library(experimentr)
library(xtable)
set.seed(5)
knitr::opts_knit$set(root.dir = tempdir())
```

```{r, echo = FALSE}
sample_user <- declare_sample(
  individuals = list(
    income = declare_variable()),
  villages = list(
    development_level = declare_variable(multinomial_probabilities = 1:5/sum(1:5))
  ),
  group_sizes_by_level = list(
    individuals = rep(1,1000), 
    villages = rep(5,200)
  ))

user_data <- make_data(sample = sample_user)

save(user_data, file = "baseline_data.RData")
```

First, we load the baseline data created by the user, and then define a set of covariates that will be simulated to conduct power analysis and for simulated analyses.

```{r covariates}
load("baseline_data.RData")

## resample (bootstrap) from user data, respecting levels
sample <- declare_sample(
  individuals = list(
    income = declare_variable()),
  villages = list(
    development_level = declare_variable(multinomial_probabilities = 1:5/sum(1:5))
  ),
  N_per_level = c(500, 10),
  data = user_data, resample = TRUE)
```
  
Second, we define the potential outcomes, which will be simulated based on the baseline covariate data.

```{r}
potential_outcomes     <-  declare_potential_outcomes(
  condition_names = c("Z0","Z1"),
  outcome_formula = Y ~ .01 + 0*Z0 + .2*Z1 + .1*income
  )
```

Fourth, we define one or more analyses we will run based on simulated data. This analysis will also be used for power analysis.


```{r}
analysis      <- declare_analysis(formula = Y ~ Z, treatment_variable = "Z", estimator = difference_in_means)
```

Then we declare the design of the experiment, in this case a simple one without clusters or blocking.

```{r}
design <- declare_design(potential_outcomes = potential_outcomes)
```

Before finalizing the design, we conduct a power analysis to determine whether 500 units and 10 clusters (villages) are sufficient. To do this, we use the diagnose function.

```{r}
simulations <- diagnose(potential_outcomes = potential_outcomes, 
                                   sample =  sample, 
                                   design = design, 
                                   analysis = analysis)
print(xtable(summary(simulations)))
```

This indicates the power, `summary(simulations)[3]`, is less than the typical threshold of `0.8`. To address this problem, we can conduct a power analysis varying the number of units in the study to find an appropriate sample size. This can be accomplished with compare_experiments.

```{r, results = 'verbatim'}
comparisons <- compare_experiments(N_per_level = list(c(1000, 20), c(500, 50), c(250, 100)),
                              potential_outcomes = potential_outcomes, 
                              sample =  sample, 
                              design = design, 
                              analysis = analysis, 
                              sims = 100)
print(xtable(summary(comparisons)))
```

After settling on a sample size and a final design, we can conduct a mock analysis of the data to ensure we are satisfied with the analysis of the data. To do this, we create mock data -- simulated from the distributions we set -- and then run the analyses on the simulated data.

```{r}
mock <- make_data(sample = sample, potential_outcomes = potential_outcomes)
mock$Z <- assign_treatment(design = design, data = mock)
mock$Y <- observed_outcome(outcome = "Y", treatment_assignment = "Z", data = mock)
```

```{r, results = 'verbatim'}
print(xtable(get_estimates(analysis = analysis, data = mock)))
```

Finally, we use the pre_register function to create the pre-registration document.


```{r, eval = FALSE}
pre <- pre_register(design = design, sample = sample, 
             potential_outcomes = potential_outcomes, analysis = analysis, 
             title = "Lady Tasting Tea", 
             author = c("Ronald A. Fisher"), 
             abstract = "Description of the lady tasting tea experiment")
```

```{r, echo = FALSE}
pre <- pre_register(design = design, sample = sample, 
             potential_outcomes = potential_outcomes, analysis = analysis, 
             title = "Lady Tasting Tea", 
             author = c("Ronald A. Fisher"), 
             abstract = "Description of the lady tasting tea experiment",
             temp_dir = TRUE, make_output = FALSE)
```
