---
title: "How to Simulate Data"
author: "Experiments in Governance and Politics"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Simulate Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,echo=FALSE}
library(experimentr)
library(knitr)
```

This vignette demonstrates how to simulate data for use in power analysis. 

# Making Covariates

## Simple User-Defined Covariates with `declare_variable()`

To make user-defined covariates, we can provide `declare_population()` 
a list of `DGP_object`s created with `declare_variable`. 
From there, we can make new random draws of the covariates either by 
supplying the `population` object to `draw_population()`, or by calling the 
`make_sample()` function that is created inside the `sample` object:

```{r}
# The sample size
N <- 10
# An unnamed, standard normal variable:
univariate_pop <- declare_population(
  declare_variable(),
  N = N
)
univariate_draw <- draw_population(population = univariate_pop)
```


```{r,echo=F}
kable(univariate_draw, digits = 2)
```

```{r}
# Multinomial, binomial and normal covariates
multivariate_pop <- declare_population(
  income = declare_variable(normal_mean = 1000,normal_sd = 10),
  gender = declare_variable(binary_probability = .5,
                            binary_categories = c("male","female")),
  region = declare_variable(multinomial_probabilities = c(.25,.25,.25,.25),
                            multinomial_categories = c("North","South","East","West")),
  employed = declare_variable(binary_probability = .7),
  N = N
)

multivariate_draw <- draw_population(population = multivariate_draw)
```


```{r,echo=F}
kable(
  multivariate_draw,
  digits = 2
  )
```

We can also generate data using custom functions, provided that a) 
the function(s) generate vectors (often randomly) of a given size, and b)
that the functions do not take arguments. 
For this reason, using 
`declare_variable()` is often more flexible, as it allows the user to change
the size of the covariate matrix.

```{r}
custom_fun_pop <- declare_population(
  # Using functions
  change_1 = function(){rnorm(n = N,mean = 0,sd = 1)},
  success_1 = function(){rbinom(n = N,size = 1,prob = .5)},
  # Using declare_variable
  change_2 = declare_variable(),
  success_2 = declare_variable(binary_probability = .5),
  N = N
)

custom_fun_draw <- draw_population(population = custom_fun_pop)
```


```{r,echo=F}
kable(
  custom_fun_draw,
  digits = 2
)
```

## Transformations of Covariates

Variables can be generated as a function of other variables using
`declare_variable()`. The user can perform a very wide range of custom 
transformations, provided that: a) the variables that are transformed have 
already been declared; b) the transformations output vectors of the right 
length; and c) the transformations take place on the same 'level' 
(more on levels below).


```{r}
# Here we create 'age' as a rounded draw from a normal distribution
transform_pop <- declare_population(
  # Declare the raw variables 
  age_raw = declare_variable(normal_mean = 30,normal_sd = 5),
  # Declare the transformation
  age_rounded = declare_variable(transformation = "round(age_raw,0)"),
  N = N
)

transform_draw <- draw_population(population = transform_pop)
```

```{r,echo=F}
kable(
  transform_draw,
  digits = 2
)
```

```{r,echo=T}
# Here use two transformations to create an 'index' of three other variables
index_pop <- declare_population(
  # Declare the raw variables 
  fitness = declare_variable(normal_mean = 10,normal_sd = 5),
  diet = declare_variable(normal_mean = 10,normal_sd = 5),
  stress = declare_variable(normal_mean = 10,normal_sd = 5), 
  # Declare the first transformation (means)
  health_mean = declare_variable(transformation = "rowMeans(cbind(fitness,diet,stress))"),
  # Declare the second transformation (integers or categories)
  health_index = declare_variable(transformation = "cut(health_mean,breaks = 4,labels = 1:4)"),
  N = N
)

index_draw <- draw_population(population = index_pop)
```

```{r,echo=F}
kable(
  index_draw,
  digits = 2
)
```

## Time-Series Data Structures

It is possible to create temporal data structures in long and wide format. 
However, in the case of complex temporally-dependent data structures, 
the user may find it more practical to code her own data-generating function and 
to use this. 

Below is an example of serially-correlated time-series data in wide format. 

```{r,echo=T}

# Here we generate wide-format temporal data 
time_series_pop <- declare_population(
  firm = list(
    # Firm profit is i.i.d in year 1
    year_1 = declare_variable(normal_mean = 10,normal_sd = 3),
    # Firm profit in subsequent years is a function of the group's profit
    year_2 = declare_variable(transformation = "year_1 + .01*sum(year_1) + rnorm(1)"),
    year_3 = declare_variable(transformation = "year_2 + .01*sum(year_2) + rnorm(1)"),
    year_4 = declare_variable(transformation = "year_3 + .01*sum(year_3) + rnorm(1)")),
  N = N
)

time_series_draw <- draw_population(population = time_series_pop)

```

```{r,echo=F}
kable(
  time_series_draw,
  digits = 2
)
```

Note that, if the list is named as above, the ID variable inherits this name
(here, `firm` becomes `firm_id`). We have here declared `firm` as a 'level'. 

Next, we have an example 

```{r,echo=T}

# Here we generate long-format temporal data 
long_time_pop <- declare_population(
  firm_year = list(
    baseline = function()rep(rnorm(N),each = 4),
    year_effect = declare_variable(),
    year = function()rep(1:N,4),
    firm = function()rep(1:N,each = 4),
    firm_obs = declare_variable(transformation = "baseline + year_effect")
    ),
  
  N = N*4
)

long_time_draw <- draw_population(population = long_time_pop)

```

```{r,echo=F}
kable(
  time_series_draw,
  digits = 2
)
```

## Multi-Level Covariate Structures

It is possible to declare data structures that have multi-level covariates,
such as individuals within schools within cities. 
However, *each level must be nested within the next one*. 
For this reason, 'time' cannot be thought of as a level, but should be 
declared in wide-format as above (`sample_6`).

In order to create levels, the user simply provides `declare_population()`
with (preferably named) lists of variables, one per level, and describes the
data structure using either `N_per_level` or `group_sizes_per_level`. 
`N_per_level` is the simplest way to specify the data structure: it tells 
us how many units or groups of units are at each level. For example, 
a study with 8 students, selected from 4 classrooms in 2 different schools would 
be represented as `c(students = 8,classes = 4,schools = 2)`.

```{r}

N_per_level <- c(students = 8,
                 classrooms = 4,
                 schools = 2)

# Note: level names are inhereted from the variable declarations, not 
# from the the N_per_level vector

school_pop <- declare_population(
 student = list(
   math_score = declare_variable(),
   lang_score = declare_variable()
 ),
 classroom = list(
   N_students_raw = declare_variable(normal_mean = 20,normal_sd = 1),
   N_students = declare_variable(transformation = "round(N_students_raw,0)")
 ),
 school = list(
   poor_area = declare_variable(binary_probability = .5)
 ),
 N_per_level = N_per_level
)

school_draw <- draw_population(population = school_pop)
```

```{r,echo=F}
kable(
  school_draw,
  digits = 2
)
```


However, if the user wants to specify specify structures with unequal group 
sizes, it is necessary to use `group_sizes_per_level`, which is a list of vectors
containing group sizes at each level. The first level should always be an 
$N$-length vector of ones, to represent the 'groups' at the unit level.
Thus, `N_per_level = 5`, `N = 5`, and `group_sizes_per_level = list(c(1,1,1,1,1))`
imply identical data structures. We repeat the example above, but with 
different class sizes:


```{r}
group_sizes_per_level <- list(
  students = rep(1,8),
  students_per_class = c(1,2,3,2),
  classes_per_school = c(3,1)
)

unequal_class_pop <- declare_population(
 student = list(
   math_score = declare_variable(),
   lang_score = declare_variable()
 ),
 classroom = list(
   N_students_raw = declare_variable(normal_mean = 20,normal_sd = 1),
   N_students = declare_variable(transformation = "round(N_students_raw,0)")
 ),
 school = list(
   poor_area = declare_variable(binary_probability = .5)
 ),
 group_sizes_per_level = group_sizes_per_level
)

unequal_class_draw <- draw_population(population = unequal_class_pop)

```


```{r,echo=F}
kable(
  
)
```


# Making Potential Outcomes 

## Simple Linear Outcomes

The potential outcomes that result from an experiment are created 
using `declare_potential_outcomes()`, which produces an object given 
One outcome is made for every combination of treatment condition and outcome, 
even if the experiment only reveals a subset of these. 
If you wish to simply create linear potential outcomes, you must specify 
the different conditions, and include a function that specifies how they 
are generated in a linear model.

```{r}
# Make a very simple sample frame
simple_frame <- declare_population(N = 10)

outcomes_1 <- declare_potential_outcomes(
  # Specify the conditions of the experiment
  condition_names = c("control","treatment"),
  # Here the average treatment effect is .5
  outcome_formula = outcome ~ 0*control + .5*treatment 
    )

linear_PO_1 <- draw_population(population = simple_frame)

##with(linear_PO_1, outcome_treatment - outcome_control)

```

## Simple Non-Linear Outcomes

Except for the case in which the user specifies the potential outcomes as 
changes in population-level proportions (discussed below), 
non-linear dependent variables need to be specified using a predictive 
equation or function in the `outcome_formula` argument of 
`make_potential_outcomes()`. This of course requires that you specify 
coefficients in a non-linear model, which can be difficult. 

* <font color='red'> *I think in the future we could just make a set of functions that they just
  give a formula to, and it does it for them* </font>

```{r}
# For a probit data-generating process, write a predictive probit function  
probit_outcome <- function(Z0,Z1){
  rbinom(n = length(Z0), size = 1, prob = pnorm(q = Z0*0 + Z1*.2))
}

# Insert the function into the outcome formula
probit_PO <- declare_potential_outcomes(
  condition_names = c("Z0","Z1"),
  outcome_formula = Y ~ probit_outcome(Z1,Z0)
)

# Make data
population <- declare_population(N = 10, potential_outcomes = probit_PO)
draw_population(population = simple_frame)

# For a Poisson data-generating process, write a predictive poisson function  
poisson_outcome <- function(Z0,Z1){
  rpois(n = length(Z0), lambda = exp(Z0*0 + Z1*.2))
}

# Insert the function into the outcome formula
pois_PO <- declare_potential_outcomes(
  condition_names = c("Z0","Z1"),
  outcome_formula = Y ~ poisson_outcome(Z1,Z0)
)
population <- declare_population(N = 10, potential_outcomes = pois_PO)

# Make data
draw_population(population = simple_frame)

```

## Simple Changes in Proportions

Some users may want to specify potential outcomes at the group-level, especially
when the outcome is binary or categorical. For example, we might want to 
specify that the treatment increases uptake of a service by 5\%, or we might 
want to specify that it increases the probability of voting for candidate A
by 10\%, while reducing that of voting for B and C by 6\% and 4\%, 
respectively. This is easily achieved with `declare_potential_outcomes()`, 
however, it means that we cannot specify outcomes as the function of covariates.

The simplest case is the binary one, where we simply provide a vector 
with the proportion of successes for each condition of the experiment. 

```{r,warning=FALSE}
# A simple yes-no difference
success_proportions <- c(control = .5, treatment_1 = .7, treatment_2 = .3)
# Treatment effects
c(True_ATE = success_proportions["treatment_1"]-
    success_proportions["control"],
  True_ATE = success_proportions["treatment_2"]-
    success_proportions["control"]
  )

proportion_PO_1 <- declare_potential_outcomes(
  condition_names = c("control","treatment_1","treatment_2"),
  population_proportions = success_proportions)

population <- declare_population(potential_outcomes = proportion_PO_1, N = 10)
draw_population(population = population)

```

Users can also provide more complicated proportional outcome structures.
For example, it is possible to specify $J$ different choices across $K$
treatment conditions by creating a $J \times K$ matrix whose columns sum to
one and bear the names of the treatment conditions, and whose rows bear
the names of the choices. It is also possible to attach a matrix of covariates.
Although the covariates are unrelated to the potential outcomes in the 
population-proportions setup, this may be useful in order to specify
clustered randomization schemes.

```{r}

treatment_conditions <- c("C","T1","T2","T3")
choices <- c("Party_A","Party_B","Party_C")

# Make the matrix of proportions
proportion_mat <- matrix(
  data = c(.7,.6,.3,.6,
           .2,.3,.3,.3,
           .1,.1,.4,.1),
  nrow = 3,ncol = 4,
  byrow = TRUE,
  dimnames = list(
    choices,
    treatment_conditions
    )
)

proportion_mat

proportion_PO_2 <- declare_potential_outcomes(
  population_proportions = proportion_mat,
  proportion_outcome_name = "vote_choice"
)

population <- declare_population(potential_outcomes = proportion_PO_2,N = 10)
draw_population(population = population)
# With a simple data structure for clustering 

voter_sample <- declare_population(
  N_per_level = c(10,2),
  level_ID_variables = c("voter_id","city_id"))

population <- declare_population(potential_outcomes = proportion_PO_2, N = 10)
draw_population(population = population)
```

## Outcomes as a Function of Covariates

We can also define outcomes as a function of unit-level covariates. 
A simple application might involve adding an error term. 

```{r}

# Adding error to outcomes_1
simple_frame_2 <- declare_population(
  error = declare_variable(normal_mean = 0,normal_sd = 1),
  N = 10
)

outcomes_1_2 <- declare_potential_outcomes(
  condition_names = c("control","treatment"),
  # Here we add error to the DGP of the outcome
  outcome_formula = outcome ~ 0*control + .5*treatment + error
    )

## NOT WORKING
##population <- declare_population(potential_outcomes = outcomes_1_2, N = 10) 
##linear_PO_2 <- draw_population(population = population)
##with(linear_PO_1,outcome_treatment - outcome_control)
##with(linear_PO_2,outcome_treatment - outcome_control)


```


We will use the schoole example from above (`sample_7`). 
Students' 
final exam score is a function of their math score, their language score,
the number of students in the classroom, a school-level fixed effect
and one of four treatment conditions:
control, private tutoring, financial incentives, and a combination 
of private tutoring and financial incentives. 

```{r,warning=FALSE}

draw_population(population = sample_7)

outcomes_2 <- declare_potential_outcomes(
  condition_names = c("control","tuition","money","tuitionxmoney"),
  outcome_formula = 
    exam_score ~ 13 + control * 0 + tuition * 3 + money * 2 + tuitionxmoney * 4 +
                 8 * math_score + 4 * lang_score + N_students * -.2 + 
                 I(school_id == 2) * 5
    )

## NOT WORKING
##population <- declare_population(potential_outcomes = outcomes_2, N = 10)
##draw_population(population = population)

```


## Clustered Outcomes 

It may sometimes be desirable to induce intra-cluster correlation into the 
data. The easiest way to do so is to generate a `cluster shock' that 
affects all units within the cluster. 

The ICC of a given variable is typically calculated as 

$$ \text{ICC} = \frac{\sigma_i^2}{\sigma_i^2 + \sigma_j^2} $$, 

where $\sigma_i^2$ is the variance among units and $\sigma_j^2$ is the variance
among the variable defined at the cluster level. It is possible to compare
different levels of ICC by generating "cluster shocks" that are added on to
the unit-level outcomes. 

```{r}
# Set the two ICC you would like to compare
ICC_1 <- .1
ICC_2 <- .6

# Make a clustered sample frames
cluster_frame <- declare_population(
  unit = list(
    baseline_1 = declare_variable(normal_mean = 0,normal_sd = sqrt(1-ICC_1)),
    baseline_2 = declare_variable(normal_mean = 0,normal_sd = sqrt(1-ICC_2))
  ),
  # The shocks that create the ICC go here
  cluster = list(
    cluster_shock_1 = declare_variable(normal_mean = 0,normal_sd = sqrt(ICC_1)),
    cluster_shock_2 = declare_variable(normal_mean = 0,normal_sd = sqrt(ICC_2))
  ),
  N_per_level = c(1000,50)
)

# We now specify potential outcomes that are generated by the low- and 
# high-ICC scenarios
cluster_PO_1 <- declare_potential_outcomes(
  condition_names = c("Z0","Z1"),
  outcome_formula = Y ~ baseline_1 + cluster_shock_1 + Z0*0 + Z1*.2
)
cluster_PO_2 <- declare_potential_outcomes(
  condition_names = c("Z0","Z1"),
  outcome_formula = Y ~ baseline_2 + cluster_shock_2 + Z0*0 + Z1*.2
)

## DOES NOT WORK
# Make both datasets
##population <- declare_population(potential_outcomes = cluster_PO_1, N = 10)
##data_1 <- draw_population(population = population)
##population <- declare_population(potential_outcomes = cluster_PO_2, N = 10)
##data_2 <- draw_population(population = population)
# Estimate the ICC
##estimate_ICC(variable = data_1$Y_Z0,cluster = data_1$cluster_id)
##estimate_ICC(variable = data_2$Y_Z0,cluster = data_2$cluster_id)

```

## Multiple Outcomes 

Often users will want to compare multiple outcomes in an experiment.
To create (independent) outcomes using the population-proportions approach,
we can simply provide draw_population() with a list of potential outcomes. 

```{r,warning=FALSE}
treatments <- c("T0","T1","T2")
multi_prop_PO <- list(
  declare_potential_outcomes(
    condition_names = treatments,
    population_proportions = c(.2,.4,.5),
    proportion_outcome_name = "outcome_1"
  ),
  declare_potential_outcomes(
    condition_names = treatments,
    population_proportions = c(.2,.9,.4),
    proportion_outcome_name = "outcome_2"
  ),
  declare_potential_outcomes(
    condition_names = treatments,
    population_proportions = c(.2,.9,.2),
    proportion_outcome_name = "outcome_2"
  )
)

population <- declare_population(potential_outcomes = multi_prop_PO, N = 10)

draw_population(population)

```

We can also define multiple outcomes that are a function of the same 
covariates, by supplying `draw_population()` with a list of potential outcomes
and a single sample frame. This has the advantage of inducing correlation
across outcomes within units.

```{r}
# We will use our health sample, created as sample_5
conditions <- c("T0","T1")

race_km <- declare_potential_outcomes(
  condition_names = conditions,
  outcome_formula = km ~ T0*0 + T1*5 + fitness + diet*.5 + I(as.integer(health_index)>1)*.3
)

hours_sleep <- declare_potential_outcomes(
  condition_names = conditions,
  outcome_formula = hours ~ T0*0 + T1*.3 + fitness*.01 + diet*.9 + I(as.integer(health_index)>1)*.3
)

health_outcomes_1 <- list(race_km,hours_sleep)

##population <- declare_population(potential_outcomes = health_outcomes_1, N = 10)
##draw_population(population = population)
```

## Outcomes as a Function of Other Outcomes

Users may want to specify outcomes as a function of other outcomes. 
For example, they might want to create an index of two outcomes. 
This is very straightforward: the relationship of the new outcomes to the 
preceding outcomes simply needs to be specified in the `outcome_formula`,
and the outcomes should be declared in succession. Continuing the previous
example:


```{r}
# Make a new potential outcome that is a function of the preceding two

index_outcome <- declare_potential_outcomes(
  condition_names = conditions,
  outcome_formula = new_index ~ I((hours_T0+km_T0)/2)*T0 + I((hours_T1+km_T1)/2)*T1
)

# Append it to the preceding outcome list
health_outcomes_2 <- c(health_outcomes_1, list(index_outcome))

draw_population(population = sample_5)

```


# Using Pre-Existing Data

## Potential Outcomes as a Function of Pre-Existing Data

... To be continued

## Re-Sampling from Pre-Existing Data

... To be continued















